---
title: "Holsenbeck_S_P3"
author: "Stephen Synchronicity"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
always_allow_html: yes
header-includes:
   - \usepackage{dcolumn}
output: 
  html_document: 
    self_contained: yes
    css: C:\Users\Stephen\Documents\R\win-library\3.4\neuhwk\rmarkdown\templates\DA5030\resources\bootstrap.min.css
    highlight: zenburn
    keep_md: no
    theme: spacelab
    toc: no
---
```{r setup, include=FALSE}
# Knitr Options
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=TRUE, fig.align='center', fig.height=5, fig.width=8, tidy=TRUE, tidy.opts=list(width.cutoff=80))
library(knitr)
knit_print.data.frame = function(x, ...) {
    res = paste(c("", "", kable(x)), collapse = "\n")
    asis_output(res)
}
options(scipen=12)
# Attach dependencies
rmarkdown::html_dependency_jquery()
rmarkdown::html_dependency_bootstrap("spacelab")
rmarkdown::html_dependency_jqueryui()
# Make reproducible
set.seed(1)
# Load packages
req.packages <- c("tidyverse","dplyr","htmltools","magrittr")
for (q in seq_along(req.packages)) {
  suppressPackageStartupMessages(library(req.packages[q],character.only = T))
}
# ----------------------- Tue Apr 03 10:20:17 2018 ------------------------#
# Custom Functions
findna <- function(df){
nalist <- sapply(sns,function(x){if(any(is.na(x)) & any(class(x) %in% c("factor","character"))){table(x, useNA = "always") -> natable;natable %>% prop.table -> naprop; return(list(Freq=natable,Prop=naprop)) }else if(any(is.na(x)) & any(class(x) %in% c("numeric","integer"))){
  summary(x)
}
})
nalist[sapply(nalist, is.null)] <- NULL
return(nalist)
  }
```
```{r 'Assignment',eval=F}
# This code will extract the assignment HTML and print the output formatted for this Rmd document. Set Assignment html below
# Use if assignment has blue font headers, and lists of questions
library(rvest)
library(xml2)
#font[contains(@color,'#24678d')]/ancestor::div[1]
htm <- xml2::read_html("https://da5030.weebly.com/practicum-3-f18.html") 
Q <- htm %>% rvest::html_nodes(xpath="//div[contains(@class,'paragraph')]|//h2[contains(@class,'wsite-content-title')]") %>% rvest::html_children()
tt <- data.frame(Tags=Q %>% html_name,Text=Q %>% html_text,stringsAsFactors = F)
lbls <- apply(tt,1,function(x){
  if(any(stringr::str_detect(x,"Problem"))){l <- "PrHd"}else if(any(stringr::str_detect(x,"Submission Details"))){l <- "Stop"}else if(any(stringr::str_detect(x,"ol"))){l <- "ol"}else if(any(stringr::str_detect(x,"ul"))){l <- "ul"}else {l <- NULL}
  return(l)})
lbls[sapply(lbls,is.null)] <- NULL
Qs <- vector("list", sum(lbls %in% "PrHd"))
for(i in seq_along(lbls)){
  if(lbls[[i]] == "Stop"){break}  #Storing following info in appropriate list item for question
 if(lbls[[i]] == "PrHd"){
  n <- Q[[i]] %>% str_extract("(?<=Problem\\s)\\d") %>% as.numeric
  h1 <- paste("#",rvest::html_text(Q[i]),"\n")
  Qs[[n]] <- list(h1=h1,q=NA)
 next}else if(length(rvest::html_attrs(Q[i]) %>% grepl("paragraph",.,ignore.case = T) & html_children(Q[i]) %>% html_attrs() %>% grepl("rgb\\(85",.,ignore.case = T)) > 0 & any(lbls[[i]] %in% c("ul","ol")) != T){
    Qs[[n]][['q']] <- paste("<div class='q'>",html_text(Q[i]),"</div>\n```{r  '",n,"'}\n```\n<p class='a'></p>\n\n",sep="")
 }else if(lbls[i] == "ol"|lbls[i] == "ul"){
   subQs <- htm %>% html_node(xpath=xml_path(Q[i])) %>% html_nodes(css="li") %>% html_text
   for(l in seq_along(subQs)){
    subQs[l] <- paste("<div class='q'>",subQs[l],"</div>\n```{r  '",n,letters[l],"'}\n```\n<p class='a'></p>\n\n",sep="")
   }
  Qs[[n]]$q <- vector()
  Qs[[n]]$q <- subQs
}
  
}
Qs <- lapply(Qs,function(x)x[order(names(x))])
# grep(pattern=substr(html_text(n),1,20),x=xml_parent(n),ignore.case = T)
# Qtext <- xml2::read_html("https://da5030.weebly.com/assignment-7.html") %>% rvest::html_nodes(xpath="//font[contains(@color,'#24678d')]/ancestor::div[1]/following-sibling::div[contains(@class,'paragraph')][1]")
# Q.form <- vector("list",length(Q))
# for (i in seq_along(Q)) {
#  Q.form[[i]] <- list(title=NA,Qs=NA)
#   Q.form[[i]][["title"]] <- Q[i] %>% rvest::html_node(css="font") %>% rvest::html_text() %>% paste("# ",.,"\n",sep="") 
#    if(length(Qtext)>0){
#     li <- xml2::xml_contents(Qtext[[i]]) %>% xml2::xml_children() %>% rvest::html_text()
#     for (l in seq_along(li)) {
#     Q.form[[i]][['Qs']][l] <- paste("## ",i,letters[l],"\n<div class='q'>",li[l],"\n</div>\n```{r '",i,letters[l],"'}\n```\n<p class='a'>\n</p>",sep="")
#     }
#    }else {
#      
#    }
# }
rapply(Qs,cat,sep="\n")
while(length(sapply(search() %in% c("package:rvest","package:xml2") %>% which(),FUN=function(x)detach(pos=x)))>0){
sapply(search() %in% c("package:rvest","package:xml2") %>% which(),FUN=function(x)detach(pos=x))}
```

# Problem 1 

<div class='q'>(0 pts) Download the data set Bank Marketing Data Set. Note that the data file does not contain header names; you may wish to add those. The description of each column can be found in the data set explanation. Use the bank-full.csv data set but also download bank.csv and use that for testing and algorithm development.</div>
```{r  '1a'}
```
<p class='a'></p>


<div class='q'>(0 pts) Explore the data set as you see fit and that allows you to get a sense of the data and get comfortable with it. Is there distributional skew in any of the features? Is there a need to apply a transform? </div>
```{r  '1b'}
```
<p class='a'></p>


<div class='q'>(20 pts) Build a classification model using a support vector machine that predicts if a bank customer will open a term deposit account.</div>
```{r  '1c'}
```
<p class='a'></p>


<div class='q'>(20 pts) Build another classification model using a neural network that also predicts if a bank customer will open a term deposit account.</div>
```{r  '1d'}
```
<p class='a'></p>


<div class='q'>(20 pts) Compare the accuracy of the two models based on absolute accuracy and AUC.</div>
```{r  '1e'}
```
<p class='a'></p>


# Problem 2 

<div class='q'>(0 pts) Download the data set Plant Disease Data Set. Note that the data file does not contain header names; you may wish to add those. The description of each column can be found in the data set explanation. This assignment must be completed within an R Markdown Notebook.</div>
```{r  '2a'}
```
<p class='a'></p>


<div class='q'>(0 pts) Explore the data set as you see fit and that allows you to get a sense of the data and get comfortable with it. Is there distributional skew in any of the features? Is there a need to apply a transform? </div>
```{r  '2b'}
```
<p class='a'></p>


<div class='q'>(40 pts) Use association rules to segment the data similar to what was done in H채m채l채inen, W., & Nyk채nen, M. (2008, December). Efficient discovery of statistically significant association rules. In Data Mining, 2008. ICDM'08. Eighth IEEE International Conference on (pp. 203-212). IEEE.</div>
```{r  '2c'}
```
<p class='a'></p>


<div class='q'>(+30 pts) Are there clusters in the data? Can plants be segmented into groups? Build a k-means clustering model to investigate. </div>
```{r  '2d'}
```
<p class='a'></p>


<div class='q'>(+10 pts) Visualize the clusters.</div>
```{r  '2e'}
```
<p class='a'></p>